Coupled response of stomatal and mesophyll conductance to light enhances photosynthesis of shade leaves under sunflecks.

### 2016-06-19 gm_newphyt/manuscript_gmes.Rmd
fixed An, gs, gm formatting; more edits to introduction

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, 2751 Australia
\
^2^ ARC Centre of Excellence for Translational Photosynthesis, Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, 2601 Australia

@@

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves. However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting carbon and water exchange in canopies.
* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower *A~n~*, g~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but g~s~ was similar. When light intensity was temporarily increased for shade leaves to match that of sun leaves, both g~s~ and g~m~ increased, leading to increases in *A~n~* greater than sun leaves. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments have important implications for up-scaling leaf level measurements and predicting whole canopy carbon gain. Despite exhibiting reduced photosynthetic biochemical capacity, the rapid up-regulation of g~m~ with increased light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.

@@

#Introduction
Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees. Predicting C uptake of forests usually involves up-scaling leaf level measurements to assess whole canopy function. Due to the costs and limitations of efficient light harvesting within plant canopies, not all leaves are exposed to full sun [@niinemets2010review], making simple up-scaling based on solar irradiance problematic. Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index (ratio of leaf area to ground area), creating a steep light gradient from the canopy top to bottom [@monsi2005factor]. Consequently, structural and functional properties of leaves within canopies are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance with canopy depth strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon]. To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A~n~* to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].

\
The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients. As *A~n~* has a saturating response with light and maximum rates depend, in part, on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does]. Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance of the growth environment [@evans2001photosynthetic]. Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization]. Higher rates of *A~n~* and stomatal conductance (g~s~) in sun leaves can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal]. Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co]; thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].

\
Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand]. This drawdown consists of two major resistance pathways to CO~2~ diffusion, namely, the CO~2~ diffusion from the atmosphere through stomata into intercellular air spaces (stomatal conductance, g~s~) and the pathway from the intercellular air spaces into the chloroplast of the mesophyll cells (mesophyll conductance, g~m~). Based on optimality theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Mesophyll conductance can also impose limitations on *A~n~* that can be as large as g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis], if g~m~ constrains CO~2~ supply to the chloroplast. Part of the variation in photosynthetic capacity between sun and shade leaves may arise from differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified. It is likely that leaf anatomical costs associated with minimizing the length and tortuosity of the g~m~ diffusion pathway are necessary to maintain the benefits of a high g~m~ [@hassiotou2009influence]. Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.

\
Additionally, accounting for short term light fluctuations within a canopy, via sunflecks, makes assessing shade leaf behaviour difficult. How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic]. The utilization of sunflecks in shade leaves is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks]. Mesophyll conductance has been shown to respond to environmental factors (e.g. CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear. For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid]. Anatomical parameters which regulate g~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light. The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance of the growth environment and as a potential response to transitory light availability.

\
Climate warming may also affect the physiological behaviour of leaves within a canopy. This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis]. How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for up-scaling *A~n~* for sun and shade leaves. Additionally, light-saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature]. How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients. The impacts of warming on plant physiological processes are vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.

\
In this study we use *Eucalyptus tereticornis* trees, planted outdoors in naturally sunlit climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to test whether the distribution of N and water supply and leaf physiological behaviour result in  higher photosynthetic capacity in sun leaves compared to shade leaves. We further aimed to quantify the contraints on An in sun and shade leaves arising from biochemical photosynthetic capacity and the CO~2~ diffusion pathway via stomatal and mesophyll conductance. As leaves which developed in the shade were expected to have lower biochemical photosynthetic capacity and correspondingly lower diffusive conductance than sun leaves, we predicted that sunfleck simulations would increase *A~n~* and g~s~ to values lower than that of sun leaves. We further predicted that climate warming would decrease g~s~ and leaf C gain in sun leaves more so thatn shade leaves during summer months, as increased evaporation demand from higher temperatures and irradiance lead to stomatal closure.

@@

##Whole tree chamber experimental design
Twelve *Eucalyptus tereticornis* Sm. seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, New South Wales, Australia. Each chamber was 9 m tall, which accommodated growth of trees for 15 months. A detailed description of the WTC operation and design is available in Barton et al. [-@barton2010whole] and methods for this experiment in [@drake2016does]. Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of ambient +3&deg;C (ET, Figure `r airvarfig`). The trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs at the end of the experiment. Trees were watered weekly with 70 L from March 2013 to September 2013. From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW. In February 2014 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering. Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (*n*=6 to *n*=3) for the final 3 months of the experiment. This limited amount of data is attributed to small sample sizes as well as the difficulty in generating a sufficient CO~2~ drawdown inside the leaf cuvette, due to drought conditions, needed to accurately measure g~m~.

@@

\
Once CO~2~ and water vapour flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400XT. The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL. Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min. This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and sample line gas streams for calculations. The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle. Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval. 
\
Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer [-@evans2013temperature], including the ternary corrections proposed by Farquhar & Cernusak [-@farquhar2012ternary], such that:

@@

where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively. For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al. [-@crous2012light] when calculating g~m~. Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are presented in Supporting Information Methods S1. The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of C~i~:C~a~. Once g~m~ was calculated, the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between g~m~ and leaf photosynthesis rate (*A~n~*) by:

@@

\
Prior to the gas exchange measurements, predawn ($\Psi_{pd}$) leaf water potentials (MPa) were measured for a separate set of sun and shade leaves using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA). The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$ before sunrise on the same day as gas exchange measurements. Following the gas exchange measurements, leaves used for gas exchange were immediately sampled for midday ($\Psi_l$) leaf water potentials. All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. Leaf water potentials and transpiration (E, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:

@@

Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E. The g~1~ parameter, describing plant water-use strategy, was estimated from ITE vs. VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near]. Small values of g~1~ indicate that transpiration is comparatively costly in C terms and reflecting conservative water use (increaed water-use efficiency), whereas large values imply a lower C cost and decreased water-use efficiency [@lin2015optimal].

@@

##Leaf resource distribution
Across six measurement campaigns over the 7 month period, PPFD was reduced on average by >75% in the shade (Figure `r parfig`). Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`). This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_l$, Table `r wtc3tab1`) nor transpiration rates (E, Table `r wtc3tab2`) differed between leaf types. Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`). Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`). No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~a~ or LMA either within or across leaf types (P > 0.05).

@@

\
Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment and temperatures (Table `r wtc3tab2`). Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c). Following an increase in light intensity to match high-light conditions, *A~n~* of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`). No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light. Photosynthesis within leaf types and warming treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).

##Stomatal conductance and leaf water-use efficiency
On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`). Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a). No effect of the warming treatment was detected on g~s~ within or across leaf types. Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).

@@

##Leaf carbon isotope discrimination and mesophyll conductance
The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`). Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments. The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment and temperatures (P < 0.001). Following the increase in PPFD from low to high light on the same leaf, g~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`). These measured values of g~m~ for shade leaves at high light were also equivalent or greater than those of sun leaves (Table `r wtc3tab2`). Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186). The large increases in g~m~ in shade leaves under high light likely resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b). No differences in g~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).

>>>>

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith 2751 NSW, Australia
\
^2^ ARC Centre of Excellence for Translational Photosynthesis, Plant Science Division, Research School of Biology, The Australian National University, Canberra 2601 ACT, Australia

@@

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves. However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting carbon and water exchange in canopies.
* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but *g*~s~ was similar. When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, leading to increases in *A*~n~ greater than sun leaves. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments have important implications for up-scaling leaf level measurements and predicting whole canopy carbon gain. Despite exhibiting reduced photosynthetic biochemical capacity, the rapid up-regulation of *g*~m~ with increased light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.

@@

#Introduction
Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees. Predicting C uptake of forests usually involves up-scaling leaf level measurements to assess whole canopy function. Due to the costs and limitations of efficient light harvesting within plant canopies, not all leaves are exposed to full sun [@niinemets2010review], making simple up-scaling based on solar irradiance problematic. Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index (ratio of leaf area to ground area), creating a steep light gradient from the canopy top to bottom [@monsi2005factor]. Consequently, structural and functional properties of leaves within canopies are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance with canopy depth strongly affects rates of leaf photosynthesis (*A*~n~) [@evans1995carbon]. To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A*~n~ to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].

\
The distribution of resources required for *A*~n~, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients. As *A*~n~ has a saturating response with light and maximum rates depend, in part, on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does]. Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance of the growth environment [@evans2001photosynthetic]. Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization]. Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A*~n~ if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co]; thus leaf-specific  hydraulic conductance (*K*~l~) should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects;@hubbard2001stomatal].

\
The photosynthetic rate in C~3~ plants is limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand]. There are two major resistance pathways to CO~2~ diffusion: the CO~2~ diffusion from the atmosphere through stomata into intercellular air spaces (stomatal conductance, *g*~s~) and the pathway from the intercellular air spaces into the chloroplast of the mesophyll cells (mesophyll conductance, *g*~m~). Based on optimality theory, regulation of *g*~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A*~n~ [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Mesophyll conductance can also impose limitations on *A*~n~ that can be as large as *g*~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A*~n~ [@niinemets2007photosynthesis], if *g*~m~ constrains CO~2~ supply to the chloroplast. Part of the variation in photosynthetic capacity between sun and shade leaves may arise from differences in *g*~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified. It is likely that leaf anatomical costs associated with minimizing the length and tortuosity of the *g*~m~ diffusion pathway are necessary to maintain the benefits of a high *g*~m~ [@hassiotou2009influence]. Nonetheless, like *g*~s~, *g*~m~ may also vary dynamically in response to environmental drivers [@flexas2008mesophyll]. Thus, stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A*~n~ across sun and shade leaves.

\
How shade leaves utilize sunflecks for short term carbon gain depends on the combined response time of *g*~s~ and *g*~m~ and the underlying photosynthetic machinery acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic]. The utilization of sunflecks in shade leaves is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks]. Mesophyll conductance has been shown to respond to environmental factors (e.g. CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than *g*~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear. For example, *g*~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid]. Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light. The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance of the growth environment and as a potential response to transitory light availability.

\
Climate warming may also affect the physiological behaviour of leaves within a canopy. This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis]. How these stresses affect the diffusion of CO~2~, through either *g*~s~ or *g*~m~, will have implications for up-scaling *A*~n~ for sun and shade leaves. Additionally, light-saturated rates of *A*~n~ are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature]. How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients. The impacts of warming on plant physiological processes are vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.

\
In this study we use *Eucalyptus tereticornis* trees, planted in naturally sunlit climate controlled whole-tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to test whether the distribution of N and water supply capacity via leaf-specific hydraulic conductance and leaf physiological behaviour result in  higher photosynthetic capacity in sun leaves compared to shade leaves. We further aimed to quantify the contraints on *A*~n~ in sun and shade leaves arising from photosynthetic capacity and the CO~2~ diffusion pathway via stomatal and mesophyll conductance. As leaves which developed in the shade were expected to have lower photosynthetic capacity and correspondingly lower diffusive conductance (both *g*~s~ and *g*~m~) than sun leaves, we predicted that sunfleck simulations would increase *A*~n~ and *g*~s~ to values lower than that attained by sun leaves in similar conditions. We further predicted that climate warming would decrease *g*~s~ and leaf C gain in sun leaves more so than shade leaves during summer months, as increased evaporation demand from higher temperatures and irradiance lead to stomatal closure.

@@

##Whole tree chamber experimental design
Twelve *Eucalyptus tereticornis* Sm. seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, New South Wales, Australia. Each chamber was 9 m tall, which accommodated growth of trees for 15 months. A detailed description of the WTC operation and design is available in Barton et al. [-@barton2010whole] and methods for this experiment in [@drake2016does]. Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of ambient +3&deg;C (ET, Figure `r airvarfig`). The trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs at the end of the experiment. Trees were watered weekly with 70 L from March 2013 to September 2013. From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW. In February 2014 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering. Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (*n*=6 to *n*=3) for the final 3 months of the experiment. This limited amount of data is attributed to small sample sizes as well as the difficulty in generating a sufficient CO~2~ drawdown inside the leaf cuvette, due to drought conditions, needed to accurately measure *g*~m~.

@@

\
Once CO~2~ and water vapour flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400XT. The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL. Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min. This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and sample line gas streams for calculations. The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle. Photosynthesis, *g*~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval. 
\
Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer [-@evans2013temperature], including the ternary corrections proposed by Farquhar & Cernusak [-@farquhar2012ternary], such that:

@@

where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, *g*~m~, respiration and photorespiration, respectively. For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al. [-@crous2012light] when calculating *g*~m~. Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are presented in Supporting Information Methods S1. The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of C~i~:C~a~. Once *g*~m~ was calculated, the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between *g*~m~ and leaf photosynthesis rate (*A*~n~) by:

@@

\
Prior to the gas exchange measurements, predawn ($\Psi_{pd}$) leaf water potentials (MPa) were measured for a separate set of sun and shade leaves using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA). The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$ before sunrise on the same day as gas exchange measurements. Following the gas exchange measurements, leaves used for gas exchange were immediately sampled for midday ($\Psi_l$) leaf water potentials. All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. Leaf water potentials and transpiration (E, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (*K*~l~, mmol m^-2^ s^-1^ MPa) with the equation:

@@

Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A*~n~ divided by E. The g~1~ parameter, describing plant water-use strategy, was estimated from ITE vs. VPD response curves by fitting a rearranged optimal *g*~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near]. Small values of g~1~ indicate that transpiration is comparatively costly in C terms and reflecting conservative water use (increaed water-use efficiency), whereas large values imply a lower C cost and decreased water-use efficiency [@lin2015optimal].

@@

##Leaf resource distribution
Across six measurement campaigns over the 7 month period, PPFD was reduced on average by >75% in the shade (Figure `r parfig`). Leaf-specific hydraulic conductance (*K*~l~) was similar across sun and shade leaves (Table `r wtc3tab1`). This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_l$, Table `r wtc3tab1`) nor transpiration rates (E, Table `r wtc3tab2`) differed between leaf types. Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`). Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`). No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, *K*~l~, E, N~a~ or LMA either within or across leaf types (P > 0.05).

@@

\
Mean *A*~n~ was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment and temperatures (Table `r wtc3tab2`). Additionally, leaf N~a~ was positively related to *A*~n~ across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c). Following an increase in light intensity to match high-light conditions, *A*~n~ of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`). No effect of the warming treatment was detected on *A*~n~ of sun leaves measured at high light or shade leaves at either low or high light. Photosynthesis within leaf types and warming treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).

##Stomatal conductance and leaf water-use efficiency
On average, *g*~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`). Photosynthesis was positively correlated with *g*~s~ in all leaves measured under high light conditions, however, *g*~s~ and *A*~n~ were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following increased PPFD, *g*~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a). No effect of the warming treatment was detected on *g*~s~ within or across leaf types. Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).

@@

##Leaf carbon isotope discrimination and mesophyll conductance
The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`). Carbon isotope discrimination associated with *g*~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments. The remainder $\Delta_o$ consists of the contributions of *g*~s~, respiration and photorespiration to discrimination.

\
Mean *g*~m~ was higher in sun compared to shade leaves (+27%) under their local light environment and temperatures (P < 0.001). Following the increase in PPFD from low to high light on the same leaf, *g*~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`). These measured values of *g*~m~ for shade leaves at high light were also equivalent or greater than those of sun leaves (Table `r wtc3tab2`). Proportional increases in *g*~m~ were matched by proportional increases in *A*~n~ from low to high light in shade leaves (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in *g*~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186). The large increases in *g*~m~ in shade leaves under high light likely resulted in the highest rates of *A*~n~ (Figure `r agmgsfig`b). No differences in *g*~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
